Exp 13 write-up

### 2023-07-16 writing/DGfS2023-abstract/write-up.tex
updated exp13 analysis and write-up

\begin{itemize}
\item[a)] projection is gradient, that is, it is not the case that content either projects or it doesn't, but rather content projects to varying degrees, reflecting either varying degrees of speaker commitment or varying degrees of certainty of the interpreter on the speaker's commitment to a content, and

@@

\item ``24 critical items in a total of 12 versions''
\\ 4 ps triggers x 6 lexicalizations = 24 critical items
\\ 12 versions = 3 (no conjunct v first vs second conjunct) x 2 (ps vs no ps) x 2 contexts

@@

``Thus comparing contexts which support the inference to contexts in which it has been made clear that the speaker is ignorant about the inference provides a way to distinguish a broad class of natural and invited pragmatic inferences from those that are really encoded as presuppositions, and thus have no choice but to project.'' (p.497).

@@

\item In this experiment, we will investigate the projection of the CCs of the 20 clause-embedding predicates using an acceptability rating task, similar to that of \citealt{mandelkern-etal2020}.

@@

Instead of MEAL's `support' context (which entails the CC), we use `neutral' contexts, which are compatible with the CC but don't entail it. These `neutral' contexts are the low and high prior probability facts of \citealt{degen-tonhauser-openmind}. Projection of the CC is possible regardless of whether Julian is German (low prior probability fact) or Cuban (high prior probability fact).

@@

\begin{exe}
\ex Explicit ignorance context:
\begin{xlist}
\ex Presupposition triggers: acceptable only under local accommodation, which is marked
\ex Non-ps triggers: acceptable because no ps is triggered
\end{xlist}
\ex Neutral context
\begin{xlist}
\ex Ps triggers: acceptable, since global accommodation is possible, no sensitivity to prior probability predicted
\ex Non-ps triggers: acceptable, since no ps is triggered, but inference comes about pragmatically, sensitivity to prior probability expected
\end{xlist}
\end{exe}

@@

\item Given \citealt{brst-lang11}, we expect SCF+ expressions to be maximally unacceptable in such contexts: at least {\em too} and {\em also}, perhaps also clefts (see \citealt{abusch10}).

@@

In the target stimuli, the interrogative sentences were identical to those of \citealt{degen-tonhauser-openmind,degen-tonhauser-language}: they were constructed from 20 clause-embedding predicates paired with 20 clauses, as shown in (\ref{sample}) for the predicate {\em discover} and the clause {\em Julian dances salsa}. There were two context conditions: In the `explicit ignorance' context, illustrated in (\ref{sample}a), the declarative sentence that preceded the interrogative conveyed that the speaker is ignorant about the content of the embedded clause. In the `neutral' context, illustrated in (\ref{sample}b-c), the declarative sentence conveyed a fact about the subject of the embedded clause that is compatible with the projection of the CC. These facts are identical to the low and high prior probability facts of \citealt{degen-tonhauser-openmind}. These neutral contexts are not the same, however, as the `support' context used in \citealt{mandelkern-etal2020} or \citealt{kalomoiros-schwarz2021}, which are identical to the projection inference. In our `neutral' context, conventionally coded presuppositions need to be globally accommodated.

The experiment also includes 6 filler stimuli, which were only presented in an explicit ignorance context and were the same for each participant. They include ``again'' (prejacent), ``stop'' (pre-state content), ``continue'' (pre-state content), for comparison with \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}, as well as ``too'' (prejacent), ``also'' (prejacent) and an it-cleft (existential closure over relative clause), because \citealt{abusch10} proposed that these are hard triggers. (Prosody is not controlled, however, for ``also'' and ``too'': the assumption is that the existential presupposition cannot be accommodated so these triggers, if indeed hard, should receive low acceptability ratings under any focus assignment.)

@@

Finally, the experiment included 4 control stimuli, which were used to exclude participants not attending to the task. These are expected to be rated as natural/acceptable.

Each participant saw a random set of 30 stimuli: each set contained one target utterance for each of the 20 contents (each paired with a unique clause-embedding predicates), the same 6 control stimuli, and the same 4  filler stimuli. 12 of the CCs were presented in the explicit ignorance context, 4 with a low prior probability fact, 4 with a high prior probability fact. Within-block trial order was randomized. After completing the experiment, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

The experiment started off with four practice trials (2 natural, 2 unnatural; participants are given feedback on correct and incorrect answers).

@@

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/explicit-ignorance-naturalness-by-predicate}
\caption{Mean naturalness ratings in explicit ignorance contexts, by expression. Error bars indicate 95\% bootstrapped confidence intervals. Violin plots indicate the kernel probability density of the individual participants' ratings. Fillers are given in gray, factive predicates in red.}\label{fig:}
\end{figure}

@@

Similar to Mandelkern et al?s 2020 analysis: They claim that there is a difference in the naturalness ratings in support and explicit ignorance contexts between presuppositions and non-presuppositions.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/naturalness-by-context-and-predicate}
\caption{Mean naturalness ratings, by context and expression. Error bars indicate 95\% bootstrapped confidence intervals.}\label{fig:}
\end{figure}

>>>>

\begin{itemize}
\item[a)] projection is gradient, that is, it is not the case that content either projects or it doesn't, but rather content projects to varying degrees, reflecting varying degrees of (of certainty of the interpreter about) speaker commitment, and

@@

\item ``24 critical items in a total of 12 versions''
\\ 4 ps triggers x 6 lexicalizations = 24 critical items
\\ 12 versions = 3 (no conjunct vs first vs second conjunct) x 2 (ps vs no ps) x 2 contexts

@@

``Thus comparing contexts which support the inference to contexts in which it has been made clear that the speaker is ignorant about the inference provides a way to distinguish a broad class of natural and invited pragmatic inferences from those that are really encoded as presuppositions, and thus {\bf have no choice but to project}.'' (p.497, bold-face added).

@@

\item In this experiment, we will investigate the projection of the CCs of the 20 clause-embedding predicates using an acceptability rating task that is similar to that of \citealt{mandelkern-etal2020}.

@@

\item Instead of MEAL's `support' context (which entails the CC), we use `neutral' contexts, which are compatible with the CC but don't entail it. (\citealt{brst-lang11} called such contexts neutral, contrasting them with positive, which is like their support context.) These `neutral' contexts are given by the low and high prior probability facts of \citealt{degen-tonhauser-openmind}: Regardless of whether Julian is German (low prior probability fact) or Cuban (high prior probability fact), he may or may not dance salsa.

Using these two types of `neutral' context allows us to identify whether acceptability ratings are also sensitive to the content's prior probability. It also gives us a comparison between contexts where the speaker is explicitly ignorant about the CC vs not explicitly ignorant but possibly ignorant.

@@

\begin{exe}
\ex Explicit ignorance context:
\begin{xlist}
\ex Presupposition triggers: acceptable only under local accommodation, which is marked
\ex Non-ps triggers: acceptable because no ps is triggered
\end{xlist}
\ex Neutral context
\begin{xlist}
\ex Ps triggers: acceptable, since global accommodation is possible, and ``default'' and ``unmarked'', no sensitivity to prior probability predicted
\ex Non-ps triggers: acceptable, since no ps is triggered, but inference comes about pragmatically, sensitivity to prior probability expected
\end{xlist}
\end{exe}

@@

\item Given \citealt{brst-lang11}, we expect SCF+ expressions, namely {\em too} and {\em also}, to be maximally unacceptable in such contexts.

\item \citealt{abusch10} leads us to expect that it-clefts receive low acceptability ratings since it is a hard trigger, like {\em too}.

@@

In the target stimuli, the interrogative sentences were identical to those of \citealt{degen-tonhauser-openmind,degen-tonhauser-language}: they were constructed from 20 clause-embedding predicates paired with 20 clauses, as shown in (\ref{sample}) for the predicate {\em discover} and the clause {\em Julian dances salsa}. There were two context conditions: In the `explicit ignorance' context, illustrated in (\ref{sample}a), the declarative sentence that preceded the interrogative conveyed that the speaker is ignorant about the content of the embedded clause. In the `neutral' context, illustrated in (\ref{sample}b-c), the declarative sentence conveyed a fact about the subject of the embedded clause that is compatible with the projection of the CC. These facts are identical to the low and high prior probability facts of \citealt{degen-tonhauser-openmind}.

The experiment also included 6 filler stimuli, which were only presented in an explicit ignorance context and were the same for each participant. They include ``again'' (prejacent), ``stop'' (pre-state content), ``continue'' (pre-state content), for comparison with \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}, as well as ``too'' (prejacent), ``also'' (prejacent) and an it-cleft (existential closure over relative clause), because \citealt{abusch10} proposed that these are hard triggers. (Prosody was not controlled: the assumption is that the  presuppositions of {\em too, also} and the it-cleft cannot be accommodated so these triggers, if indeed hard, should receive low acceptability ratings under any focus assignment.)

@@

Finally, the experiment included 4 control stimuli, which were used to exclude participants not attending to the task. These were expected to be rated as natural/acceptable.

Each participant saw a random set of 30 stimuli: Each set contained one target utterance for each of the 20 CCs (each paired with a unique clause-embedding predicates), the same 6 control stimuli, and the same 4  filler stimuli. 12 of the CCs were presented in the explicit ignorance context, 4 with a low prior probability fact, 4 with a high prior probability fact. Within-block trial order was randomized. After completing the experiment, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

The experiment started off with four practice trials (2 natural, 2 unnatural; participants were given feedback on correct and incorrect answers).

@@

Data points:

\item Each predicate and each CC received 370 ratings.

\item Each of the 60 predicate/context combinations received at least 59 ratings (median: 80, mean: 123).

\item Each predicate received at least 200 ratings in the explicit ignorance context (median: 224.5, mean: 222).

\item Each predicate received at least 59 ratings in each of the neutral contexts (median: 75, mean: 74).

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/explicit-ignorance-naturalness-by-predicate}
\caption{Mean naturalness ratings in explicit ignorance contexts by expression. Error bars indicate 95\% bootstrapped confidence intervals. Violin plots indicate the kernel probability density of the individual participants' ratings. Fillers are given in gray, factive predicates in red.}\label{fig:acc-by-expression}
\end{figure}

@@

\citealt{mandelkern-etal2020} did not merely compare acceptability ratings of presuppositions and non-presuppositions in explicit ignorance contexts, but they rather claimed that there is a difference in the acceptability ratings in the two contexts between presuppositions and non-presuppositions. Specifically, while there is no difference for non-presuppositions, the ratings for presuppositions are lower in the explicit ignorance context than in the support context.

For comparison with their way of presenting the results, Fig.~\ref{fig:acc-by-context2} plots the mean acceptability ratings by expression and context (distinguishing explicit ignorance from neutral), with the expressions sorted by the difference between the mean rating in the two contexts: The mean rating in the neutral context is higher than the mean rating in the explicit ignorance context for {\em be annoyed}; the opposite is the case for {\em confirm}.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/naturalness-by-context2-and-predicate}
\caption{Mean naturalness ratings by expression and context. Error bars indicate 95\% bootstrapped confidence intervals.}\label{fig:acc-by-context2}
\end{figure}

\item For {\em be annoyed} and {\em know}, the mean rating in the explicit ignorance context is lower than the mean rating in the neutral context, suggesting that the interrogatives are judged to be less neutral when the speaker is explicitly ignorant about the truth of the CC.

For {\em see}, the mean rating in the two contexts does not appear to be different (but model suggests neutral is already higher), and the pattern seems to go in the opposite direction for {\em discover} and {\em reveal}.

Thus, on this way of displaying the data, too, factive predicates do not show a homogenous pattern that distinguishes them from non-factive predicates.

\item Non-factive predicates also do not show a homogenous pattern: For some predicates, the ratings in the two contexts do not appear to be different (e.g., {\em acknowledge, inform}), while ratings in the explicit ignorance context seem to be higher than in the neutral context for other predicates (e.g., {\em say, confirm}).

\item When ratings are higher in explicit ignorance than neutral contexts, this suggests that a natural use of these predicates is for the speaker to inquire about the CC, given their explicit ignorance.

\subsection{The effect of the prior probability of the context on naturalness ratings}

Fig.~\ref{fig:acc-by-context} plots the mean acceptability ratings by expression and context (distinguishing explicit ignorance from factL and factH), with the expressions sorted on the mean acceptability rating in the explicit ignorance context, as in Fig.~\ref{fig:acc-by-expression}.

\begin{figure}[h!]
\centering
\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/naturalness-by-context-and-predicate}
\caption{Mean naturalness ratings by expression and context. Error bars indicate 95\% bootstrapped confidence intervals.}\label{fig:acc-by-context}
\end{figure}

\item For all predicates except {\em pretend} (and {\em admit}?), the acceptability ratings are higher in factH contexts than in factL contexts. This observation suggests that acceptability ratings are sensitive to the prior probability of the CC. This is true regardless of whether ratings in explicit ignorance contexts are higher or lower or similar to those in neutral contexts.

This observation suggests that it is generally more acceptable for a speaker to ask about some AH's relation to the CC in a context in which the speaker assigns a higher probability to the CC than in a context where the speaker assigns a lower probability to the CC.

\item {\em be annoyed} and {\em know} show distinct patterns: For {\em be annoyed}, ratings are higher in the factH context than in the factL context, and higher in the factL context than in the explicit ignorance context. For {\em know}, the latter two ratings do not differ. So even these two predicates don't show the same pattern in this task.

\item Assuming that both predicates are factive and trigger a ps doesn't capture why they show these different patterns.

\item {\em be annoyed:} According to \citealt{karttunen2016}, this predicate codes that the AH believes the CC. Belief of the speaker in the CC arises as long as there is no contextual evidence/information that would justify why the speaker wouldn't share the belief. In the explicit ignorance context, the speaker has gone on record that they are ignorant about the CC, that is, do not share in the AH's belief. Given the assumption that {\em be annoyed} only codes AH belief in the CC, the very low acceptability ratings in explicit ignorance contexts are not accounted for.

\item An explanation may have to do with at-issueness: The explicit ignorance context makes the question ?CC at-issue, but  the interrogative makes the CC not-at-issue, resulting in lower acceptability ratings.

\item The same explanation may also work for the fillers with {\em stop} and {\em continue}: The explicit ignorance context makes the pre-state content at-issue, but the interrogative makes it not-at-issue.

\item The CC of {\em know} is still quite not-at-issue, but not as much as that of {\em be annoyed}. {\bf need to check this in our other data}

\item For predicates other than {\em be annoyed} and {\em know}, acceptability ratings in explicit ignorance contexts are roughly comparable to those in factH contexts. That is, asking whether an AH has a relation to the CC is just as acceptable in a context in which the speaker is explicitly ignorant of the CC and in a context in which the speaker has a belief relative to which the CC has a high probability.

It is when the speaker has a belief relative to which the CC has a low probability that the acceptability ratings are lowest (except with {\em pretend}).
